---
output: github_document
---
# generating LiDAR metrics with the lidR package.

The code used in these examples is taken from the following publication: 

Michael J Mahoney, Lucas K Johnson, Eddie Bevilacqua & Colin M Beier
(2022) Filtering ground noise from LiDAR returns produces inferior models of
forest aboveground biomass in heterogenous landscapes, GIScience & Remote
Sensing, 59:1, 1266-1280, DOI: 10.1080/15481603.2022.2103069

## Example:

First, we load the required libraries and the source functions. The 
`lidar-metrics.R` file contains the functions that are used to generate the 
metrics. 
```{r}
suppressPackageStartupMessages({
  library(lidR)
  library(proceduralnames) # used in the source functions.
  library(terra)
  library(sf)
})

source("lidar-metrics.R")
```

Then we load up some example data and of a point cloud of trees and create some
fake plots to extract metrics from.
```{r}
# load an example dataset from the lidR package
LASfile <- system.file("extdata", "MixedConifer.laz", package = "lidR")
trees <- readLAS(LASfile, filter = "-drop_z_below 0")

# plot(trees) to view the point cloud data in 3D

# make some fake plots:
tree_crs <- st_crs(trees)
trees_bbox <- lidR::st_bbox(trees) |>
  sf::st_as_sfc()
trees_cent <- st_centroid(trees_bbox)
plot_max_ext <- (trees_bbox - trees_cent) * 0.75 + trees_cent
fake_plots <- sf::st_sample(x = plot_max_ext, size = 9, type = "regular") |>
  sf::st_buffer(5) |>
  sf::st_set_crs(tree_crs)

plot(fake_plots, col = "#cc4040c9")
```

The `compute_pixel_metrics` function is a wrapper around 
`lidR::pixel_metrics` that generates 40 metrics from the point cloud data and 
saves them to a raster file. 

```{r warning=FALSE}
# This uses the `compute_pixel_metrics` which is a wrapper around
# `lidR::pixel_metrics`. It generates 40 rastersied metrics from the
# point cloud data.
gridded_metrics <- compute_pixel_metrics(
  trees,
  res = 2,
  output_filename = "grid_metrics.tif"
)
gm <- terra::rast(gridded_metrics)
print(names(gm))
plot(gm[[1:12]], col = hcl.colors(100, "viridis"))
```

The `compute_plot_metrics` function is a wrapper around `lidR::plot_metrics` 
that generates metrics for a set of polygons. The `rh_preds` function is a 
reduced/modified version of the `lidar_preds` function in the "lidar-metrics.R" 
script that can be used with `lidR::plot_metrics` to generate relative height 
metrics at 1 percentile intervals.
```{r warning=FALSE}
# now to extract metrics for these plots we use `compute`
plot_metrics_sf <- compute_plot_metrics(trees, geometry = fake_plots) |>
  sf::st_as_sf()
print(plot_metrics_sf)

# or if you just want relative height metrics at 1 percentile intervals you
# can use:
rh_mets_sf <- lidR::plot_metrics(
  trees,
  func = rh_preds(Z),
  geometry = fake_plots
) |>
  sf::st_as_sf()
print(rh_mets_sf)
# the rh_preds function is just a reduced/modified version of the lidar_preds
# these examples should demonstrate how to build any custom functions that
# can be used with `lidR::plot_metrics` or `lidR::pixel_metrics`


plot(gm[[2]], col = hcl.colors(100, "viridis"))
plot(rh_mets_sf["rh99"], add = TRUE, key.pos = 1, axes = TRUE, axes = TRUE)
```

